---
title: "Reproducible Research: Peer Assessment 1"
author: "Mdcrab02"
date: "January 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load the data
```{r load}
data <- read.csv("activity.csv", header=TRUE)
```

##Preprocess the data.  Change date to datetime type.
```{r preprocess}
data$date <- as.Date(data$date, "%Y-%m-%d")
```

##Make a histogram of the total number of steps taken each day
``` {r histogram}
data2 <- aggregate(steps ~ date, data, sum)
hist(data2$steps)
``` 


##What is mean total number of steps taken per day?
``` {r average}
cat("mean:", mean(data2$steps))
cat("median:",median(data2$steps))
```

##What is the average daily activity pattern?
``` {r patterns}
data_5 <- aggregate(steps ~ interval, data, mean, na.action = na.omit)
plot(steps ~ interval, data_5, type="l", main="Average number of steps taken", sub="5-minute interval, averaged across all days")
``` 


Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r maximum}
data_5[data_5$steps == max(data_5$steps),]
``` 

##Imputing missing values
Calculate the total number of missing values in the dataset (i.e. the total number of rows with NAs).

``` {r missing}
sum(!complete.cases(data))

summary(data)
``` 

Filling in all of the missing values in the dataset.
Using the mean for that 5-minute interval.
``` {r filling}
missing <- data[!complete.cases(data),]

temp <- merge(missing, data_5, by="interval",all.x=T)
colnames(temp) <- c("interval", "tmp", "date", "steps")

new_data <- rbind(data[complete.cases(data),c("steps", "date", "interval")], temp[,c("steps", "date", "interval")])
``` 

A histogram of the total number of steps taken each day. 
``` {r new_histogram}
new_data2 <- aggregate(steps ~ date, new_data, sum)
hist(new_data2$steps)
``` 

Calculate the mean and median total number of steps taken per day.
``` {r new_average}
cat("mean:", mean(new_data2$steps))
cat("median:",median(new_data2$steps))
``` 

Are these values different from the estimates from the first part of the assignment?
``` {r clash_of_averages}
#Average
if (mean(new_data2$steps) != mean(data2$steps)) {
  print("mean changed")
} else {
  print("mean did not change")
}
#Median
if (median(new_data2$steps) != median(data2$steps)) {
  print("median changed")
} else {
  print("median did not change")
}
``` 

What is the impact of imputing missing data on the estimates of the total daily number of steps?
``` {r impact}
cat("Imputing missing data changed the sum of total daily number of steps by:",sum(new_data2$steps, na.rm = T) - sum(data2$steps, na.rm = T))
``` 




## Are there differences in activity patterns between weekdays and weekends?
``` {r difference}
require(ggplot2)

new_data$weekend <- ifelse(weekdays(new_data$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")

data_weekend <- aggregate(steps ~ interval + weekend, new_data, mean, na.action = na.omit)

ggplot(data=data_weekend, aes(x=interval, y=steps, color=weekend), main="Average number of steps taken") + facet_wrap(~weekend) + geom_line()

``` 